Learning Action Models from State Observations

### 2018-01-26 doc/ijcai-18/observations-learning.tex
Add files via upload
Section 1 included

\section{Introduction}
Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}. At these planning tasks, automated planners are required to reason about an action model that correctly and completely captures the possible world transitions~\cite{geffner:book:2013}. Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of {\em AI planning}~\cite{kambhampati:modellite:AAAI2007}.

The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc AMAN}~\cite{zhuo2013action}, {\sc ARMS}~\cite{yang2007learning}, {\sc LOCM}~\cite{cresswell2013acquiring} or {\sc SLAF}~\cite{amir:alearning:JAIR08}. Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper presents a novel approach for learning \strips\ action models that introduces the following contributions:

\begin{enumerate}
\item {\em Learning action models as planning}.  The practicality of the compilation approach allow us to report results over a wide range of planning domains. Furthermore, opens up a way towards the {\em bootstrapping} of planning action models (a planner exploring its state space and using the obtained data to learn/update its action model).
\item {\em Learning from state observations}.  Our approach does not require the learning samples to be observations of actions but simply state observations, broadening the range of application to external observers.
\item {\em Model evaluation}. The compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}. 
\end{enumerate}

>>>>

\section{Introduction}

The automated learning of action models in planning is a promising approach, as an alternative to hand-coding by a human or when the action model is not available. Action-model acquisition has been addressed with various algorithms. {\sc ARMS}~\cite{yang2007learning} receives  plan traces as training samples and defines the learning task as a set of weighted constraints that must hold for the plans to be correct, which is then solved by a MAX-SAT solver. {\sc ARMS} validates the learned model with a test set of example plans of six domains from the IPC %checking missing preconditions and redundant effects.
The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} generates a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the true (reference) model. In the case of {\sc LOCM}~\cite{cresswell2013acquiring}, the learning samples are only the examples plans, without providing information about predicates or states. {\sc LOCM} relies on assumptions on the kind of domain structure of the model and it was able to correctly discover the dynamics of the domain in 12 out of 20 tested domains by comparison with a reference model \cite{GregoryC16}. Finally, {\sc AMAN}~\cite{zhuo2013action} is able to work with incorrectly observed plan traces and the accuracy of the learned models is validated in three IPC domains with respect to the reference model.

The learning samples required by all the aforementioned approaches are observations of plan executions in the form of plan traces. A more recent work proposes acquiring symbolic action models from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}. Additionally, all the above approaches but {\sc ARMS} validate the accuracy of the learned model with respect to a reference model.

Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models. The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemas (the learned model). \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model. The practicality of the compilation approach allow us to report results over a wide range of 15 IPC planning domains.

As for the type of training samples, we adopt a middle way between unstructured inputs and plan traces, wherein only samples that are state observations are required, without further information about state transitions. \textbf{Learning from state observations} is a relevant advancement as in many applications (e.g. computer games) the actual actions executed by the agent are not directly observable but instead we can observe the resulting states. Handling state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition tasks \cite{SohrabiRU16}.

We propose a two-fold model validation. The first method compares the learned model to a reference model, evaluating the syntactic and semantic similarity between both. In the second method, the learned model is tested with a set of state observation sequences. Since plan traces are not available, we extend the compilation to accept a learned model as input besides the observations. This new task consists in learning a new model that induces the observations whilst measuring the \emph{edit distance} to the input model. This way, the compilation scheme is also extensible to assess how well a \strips\ action model matches a given set of observations and it can thus be regarded as a first step to apply \textbf{action-model recognition}.

This paper is organized as follows ...
